Temporal Semantics to Go: A Complete Run of the FraCaS Test Suite

### 2020-12-19 paper2020/paper_2021.tex
IWCS_format

\newif\ifanon\anonfalse

\documentclass[a4paper]{article}

@@

\begin{document}

\title{FraCaS: Temporal Analysis}
% \author{\ifanon Anonymous for review\else
%   \authorname{Jean-Philippe Bernardy} \authorname{Stergios Chatzikyriakidis}
%   \affiliation{University of Gothenburg}
%   \email{firstname.lastname@gu.se}\fi}

\author{\ifanon Anonymous for review\else
  {Jean-Philippe Bernardy}\footnote{University of Gothenburg} ~~ {Stergios Chatzikyriakidis}\footnote{University of Gothenburg}
}

\maketitle

@@

\abstract{In this paper, we propose an implementation of temporal semantics
  which is suitable for inference problems. This implementation
  translates syntax trees to logical formulas, suitable for
  consumption by the Coq proof assistant. We support several phenomena
  including: temporal references, temporal adverbs, aspectual classes
  and progressives.  We apply these semantics to the complete FraCaS testsuite.
  We obtain an accuracy of 81 percent overall and 73 percent for problems explicitly marked
  as related to temporal reference.
}

@@

\paragraph{Dynamic Semantics}
Parse trees are then processed by a dynamic semantic component.
Its role is essentially to support  (non-temporal) anaphora, using a
monadic-based dynamic semantics, generally following the state of the
art in this matter
\citep{unger_dynamic_2011,charlow_monadic_2015,charlow_modular_2017}.
Our particular implementation does not handle every case \ifanon
(citations anonymized) \else \citet{bernardy_computational_2020,bernardy_wide-coverage_2019} \fi
but non-temporal anaphoroi in the testsuite are generally resolved as
they should be: on the whole accuracy is not affected significantly by
issues in this subsystem.

@@

We support additionally the basic constructions and phenomena present
in the testsuite, including adjectives, adverbs, nouns, verbs,
anaphora, etc. The method is outlined by
\citet{montague_english_1970,montague_proper_1973}, but we direct the
reader to our previous work for details \ifanon (citations anonymized)
\else \citet{bernardy_type-theoretical_2017,bernardy_wide-coverage_2019} \fi, but the particular
treatment of such phenomena is essentially independent from our
treatment of time: in this paper we simply ignore these aspects beyond
the fact that they are handled correctly in most of the cases of the
FraCaS testsuite.

@@

\begin{equation}
  \label{eq:smith-novel-1}
\begin{array}{l}
∀x. novel(x) ∧ \\
∃t_1,t_2. [t_1,t_2] ⊆ 1991 ∧ write(smith,x,t_1,t_2) ∧ \\
∃t_3,t_4. [t_3,t_4] ⊆ 1992 ∧ write(smith,x,t_3,t_4) \\
⟶ ⊥ \\
\end{array}
\end{equation}
Note here that the scope for the existential is extended beyond the
scope of \textbf{P1}, and its polarity switched (to universal). This extension can follow the account of
\citet{unger_dynamic_2011}, and our implemented analysis of
anaphora\ifanon (citation anonymized)
\else\citep{bernardy_computational_2018,bernardy_wide-coverage_2019}\fi.

@@

Our dynamic and temporal semantics gives the following interpretation
for \verb!s_279_1_p! implies \verb!s_279_3_h!.
{\small
\begin{verbatim}
FORALL (fun a=>novel_N a)
(fun a=>(exists (b: Time),
((exists (c: Time),
(IS_INTERVAL Date_19910101 b /\
 IS_INTERVAL c Date_19911231 /\
 IS_INTERVAL b c /\
 appTime b c (write_V2 a)
   (PN2object smith_PN))))) ->
Not (exists (f: Time),
((exists (g: Time),
 (IS_INTERVAL Date_19920101 f /\
  IS_INTERVAL g Date_19921231 /\
  IS_INTERVAL f g /\
  appTime f g (write_V2 a) 
    (PN2object smith_PN)))))).
\end{verbatim}
}
In the above, one should remark the top-level quantification over the
novel (as explained in \cref{sec:our-semantics}), the quantification
over time intervals as individual timepoints, and the use of custom
operators for several constructions (\verb!FORALL!, \verb!Not!, \verb!IS_INTERVAL!,
\verb!appTime!). This use of custom operators is useful for several
generalisations (for example, we have quantifiers such as \verb!MOST! in
addition to \verb!FORALL! --- see \ifanon (citation anonymized)
\else \citet{bernardy_type-theoretical_2017} \fi)

@@

The intros and destruct tactics serve bookkeeping purposes. The
critical part is the use of the \texttt{writeUnique} axiom, which
witnesses the aspectual class of the predicate \texttt{write\_V2}.
The proof is completed by the use of the \texttt{lia} tactic, which is
embeds a decision procedure for linear arithmetic
problems. Fortunately, \texttt{lia} can take care of all the problems
which arise in the FraCaS testsuite.

\begin{figure*}
\begin{subfigure}{0.47\textwidth}

@@

  \caption{Selected correctly handled examples}
  \label{fig:correct-examples}
\end{subfigure}
\hfill
\begin{subfigure}{0.47\textwidth}
{
} \fracasex{259	}{\answer yes
\item	The conference started on July 4th, 1994.
\item	It lasted 2 days.
\hyp    The conference was over on July 8th, 1994.
\failsf Incorrect lexical semantics for ``last $n$ days''.
} \fracasex{294	}{\answer yes
\item	Smith was running his own business in two years.
\hyp    Smith ran his own business.
\failsf Parsing error for ``in two years''
} \fracasex{317	}{\answer yes
\item	Every representative has read this report.
\item	No two representatives have read it at the same time.
\item	No representative took less than half a day to read the report.
\item	There are sixteen representatives.
\hyp    It took the representatives more than a week to read the report.
\failsf Counting semantics not implemented
% } \fracasex{318	}{\answer no
% \item	While Jones was updating the program, Mary came in and told him about the board meeting.
% \item	She finished before he did.
% \hyp    Mary's story lasted as long as Jones's updating the program.
% \failsf Incorrect interaction with anaphora
}
  \caption{Selected incorrectly handled examples.
  We list below a number of selected examples from the temporal section
of the FraCaS testsuite where our semantics fails to list the correct
inference. We list the reason why it fails under the item \textbf{F}.
}
  \label{fig:incorrect-examples}
\end{subfigure}
% \caption{Selected examples.}
\end{figure*}

\section{Results and Evaluation}
\label{sec:results}
Our target is the FraCaS testsuite, which aims at covering a wide
range of common natural-language phenomena. 
%
The suite is structured according to the semantic phenomena involved
in the inference process for each example, and contains nine sections:
Quantifiers, Plurals, Anaphora, Ellipsis, Adjectives, Comparatives,
Temporal, Verbs and Attitudes.  The system described here focuses on
the Temporal section. However, it also supports the other eight
sections.  To our knowledge this is the first system which attempts to
target the temporal section in full. But in fact, our system even provides
support for all the other sections. Thus, a couple of decades after its
formulation, we propose a first attempt at covering the whole suite.
As such, there it is no other system to compare our system with, in
all aspects.  We can however compare with systems which target parts
of the FraCaS testsuite, as shown in \cref{tab:results}.

\providecommand\ncases[1]{{\ensuremath{^{#1}}}}
\begin{table}[hbt]
  \centering
  \small
\begin{tabularx}{\columnwidth}{Xr@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r}
Section      & {\kern -2em} \#FraCaS
                          & This        & FC2         & FC & MINE & Nut  & LP  \\ \hline
Quantifiers  & 75         & .93        & .96         & .96    & .77  & .53  & .93  \\
             &            & \ncases{74}& \ncases{74} &        &      &      &     \ncases{44} \\
Plurals      & 33         & .79        & .82         & .76    & .67  & .52  & .73 \\
             &            &            &             &     &   &   & \ncases{24} \\
Anaphora     & 28         & .79        & .86         &   -    & -    & -    &  -       \\
Ellipsis     & 52         & .81        & .87         &   -    & -    & -    &  -       \\
Adjectives   & 22         & .95        & .95         & .95    & .68  & .32  & .73 \\
             &            & \ncases{20}&  \ncases{20}&     &   &   &  \ncases{12} \\
Comparatives & 31         & .65        & .87         & .56    & .48  & .45  &  -       \\
Temporal     & 75         & .73        &  -          &   -    &   -  &  -   &  -       \\
Verbs        & 8          & .75        & .75         &   -    & -    & -    &  -       \\
Attitudes    & 13         & .85        & .92         & .85    & .77  & .46  & .92  \\ 
             &            &            &             &        &      &      & \ncases {9}  \\ \hline
Total        & 337        & .81        & .89         & .83    & .69  & .50  & .85  \\
             &            & \ncases{329}& \ncases{259}& \ncases{174}  & \ncases{174}& \ncases{174}& \ncases{89}
  \end{tabularx}
  \caption{Accuracy of our system compared to others.
    ``This" refers to the approach presented in this paper. When a
    system does not handle the nominal number of test cases (shown in
    the second column), the actual number of test cases attempted is
    shown below the accuracy figure, in smaller font.  ``FC''
    refers to the work of \citet{bernardy_type-theoretical_2017}, and ``FC2'' its followup~\citep{bernardy_wide-coverage_2019}. ``MINE" refers
    to the approach of \citet{Mineshima:2015}, ``NUT" to the CCG
    system that utilises the first-order automated theorem prover
    \textit{nutcracker} \cite{bos:2008}, and ``Langpro"
    to the system presented by \citet{abzianidze_tableau_2015}. A dash
    indicates that no attempt was made for the section. }
  \label{tab:results}
\end{table}
Additionally, to provide the reader with a better sense of what the
complete system is capable and not capable of, we list a number of
correctly and incorrectly handled examples in \cref{fig:correct-examples} and \cref{fig:incorrect-examples}.
\paragraph{Interaction with anaphora}
\label{sec:interaction-with-anaphora}
One reason explaining the lower performance of our system on some
sections of the testsuite is that our interpretation of time interacts
imperfectly with anaphora and ellipsis. Consider the following example:

@@

We have presented a first attempt for a computational approach dealing
with the temporal section of the FraCaS test suite. To do this, we
have provided a new, simplified taxonomy of aspectual classes for verb
phrases, guided by the applicability of the unicity of action and
temporal subsumption properties. While part of this simplification is
accidental (conflation of activity and accomplishment), we find that
other parts (the automatic distinction between repeatable and
unrepeatable achievements) constitute theoretical improvements.

@@

\cite{bernardy_wide-coverage_2019} presented a
logical system for handling 8 of the 9 sections of the FRACAS test
suite, but excluded section 7, suggesting that it requires many
examples that need an \textit{ad hoc} treatment. Here, we take up this
challenge and show that a system similar to theirs can be extended to
cover the remaining section of the test suite.
%
We use the same combination of a number of well-studied tools: type
theory, parsing using the Grammatical Framework (GF), Monadic Dynamic
Semantics and proof assistant technology (Coq). The system achieves an
accuracy of 0.73 on the Temporal Section and 0.81 overall.
One of the
things to be looked at is fixing the issues associated with parts of
the test suite that ``broke'' when the temporal analysis was
introduced. Some of these have been already mentioned: interaction of
the temporal variables with anaphora.

@@

\section*{Acknowledgements}

\ifanon
Anonymized
\else
The research reported in this paper was supported by grant 2014-39 from the
Swedish Research Council, which funds the Centre for Linguistic Theory and
Studies in Probability (CLASP) in the Department of Philosophy, Linguistics,
and Theory of Science at the University of Gothenburg. We are grateful to
our colleagues in CLASP for helpful discussion of some of the ideas presented
here. We also thank anonymous reviewers for their useful comments on an
earlier draft of the paper.
\fi
\bibliographystyle{apalike}

>>>>

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{acl2020}
\usepackage{times}
\usepackage{latexsym}
\renewcommand{\UrlFont}{\ttfamily\small}

@@

%\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\newcommand\BibTeX{B\textsc{ib}\TeX}

\title{Temporal Semantics to Go: A Complete Run of the FraCaS Test Suite}

\author{First Author \\
	Affiliation / Address line 1 \\
	Affiliation / Address line 2 \\
	Affiliation / Address line 3 \\
	\texttt{email@domain} \\\And
	Second Author \\
	Affiliation / Address line 1 \\
	Affiliation / Address line 2 \\
	Affiliation / Address line 3 \\
	\texttt{email@domain} \\}

\date{}

\begin{document}
	\maketitle

@@

\begin{abstract}In this paper, we propose an implementation of temporal semantics that
  translates syntax trees to logical formulas, suitable for
 consumption by the Coq proof assistant. The analysis  supports a wide range of phenomena
 including: temporal references, temporal adverbs, aspectual classes
 and progressives.  The new semantics are built on top of a previous system handling all sections of the FraCaS test suite except the temporal reference section, and we  obtain an accuracy of 81 percent overall and 73 percent for the problems explicitly marked
as related to temporal reference. To the best of our knowledge, this is the best performance of a logical system on the whole of the FraCaS. 
\end{abstract}

@@

\paragraph{Dynamic Semantics}
Parse trees are then processed by a dynamic semantic component.
Its role is essentially to support  (non-temporal) anaphora, using a
monadic-based dynamic semantics, generally following the state of the
art in this matter
\citep{unger_dynamic_2011,charlow_monadic_2015,charlow_modular_2017}.
Our particular implementation does not handle every case 
(citations anonymized) \citet{bernardy_computational_2020,bernardy_wide-coverage_2019} \fi
but non-temporal anaphoroi in the testsuite are generally resolved as
they should be: on the whole accuracy is not affected significantly by
issues in this subsystem.

@@

We support additionally the basic constructions and phenomena present
in the testsuite, including adjectives, adverbs, nouns, verbs,
anaphora, etc. The method is outlined by
\citet{montague_english_1970,montague_proper_1973}, but we direct the
reader to our previous work for details  (citations anonymized)
\citet{bernardy_type-theoretical_2017,bernardy_wide-coverage_2019}, but the particular
treatment of such phenomena is essentially independent from our
treatment of time: in this paper we simply ignore these aspects beyond
the fact that they are handled correctly in most of the cases of the
FraCaS testsuite.

@@

\begin{equation}
  \label{eq:smith-novel-1}
\begin{array}{l}
∀x. novel(x) ∧ \\
∃t_1,t_2. [t_1,t_2] ⊆ 1991 ∧ write(smith,x,t_1,t_2) ∧ \\
∃t_3,t_4. [t_3,t_4] ⊆ 1992 ∧ write(smith,x,t_3,t_4) \\
⟶ ⊥ \\
\end{array}
\end{equation}
Note here that the scope for the existential is extended beyond the
scope of \textbf{P1}, and its polarity switched (to universal). This extension can follow the account of
\citet{unger_dynamic_2011}, and our implemented analysis of
anaphora (citation anonymized)
\citep{bernardy_computational_2018,bernardy_wide-coverage_2019}.

@@

Our dynamic and temporal semantics gives the following interpretation
for \verb!s_279_1_p! implies \verb!s_279_3_h!.
{\small
\begin{verbatim}
FORALL (fun a=>novel_N a)
(fun a=>(exists (b: Time),
((exists (c: Time),
(IS_INTERVAL Date_19910101 b /\
 IS_INTERVAL c Date_19911231 /\
 IS_INTERVAL b c /\
 appTime b c (write_V2 a)
   (PN2object smith_PN))))) ->
Not (exists (f: Time),
((exists (g: Time),
 (IS_INTERVAL Date_19920101 f /\
  IS_INTERVAL g Date_19921231 /\
  IS_INTERVAL f g /\
  appTime f g (write_V2 a) 
    (PN2object smith_PN)))))).
\end{verbatim}
}
In the above, one should remark the top-level quantification over the
novel (as explained in \cref{sec:our-semantics}), the quantification
over time intervals as individual timepoints, and the use of custom
operators for several constructions (\verb!FORALL!, \verb!Not!, \verb!IS_INTERVAL!,
\verb!appTime!). This use of custom operators is useful for several
generalisations (for example, we have quantifiers such as \verb!MOST! in
addition to \verb!FORALL! --- see (citation anonymized)
 \citet{bernardy_type-theoretical_2017} )

@@

The intros and destruct tactics serve bookkeeping purposes. The
critical part is the use of the \texttt{writeUnique} axiom, which
witnesses the aspectual class of the predicate \texttt{write\_V2}.
The proof is completed by the use of the \texttt{lia} tactic, which is
embeds a decision procedure for linear arithmetic
problems. Fortunately, \texttt{lia} can take care of all the problems
which arise in the FraCaS testsuite.
\iffalse
\begin{figure*}
\begin{subfigure}{0.47\textwidth}

@@

  \caption{Selected correctly handled examples}
  \label{fig:correct-examples}
\end{subfigure}
\hfill
\begin{subfigure}{0.47\textwidth}
{
} \fracasex{259	}{\answer yes
\item	The conference started on July 4th, 1994.
\item	It lasted 2 days.
\hyp    The conference was over on July 8th, 1994.
\failsf Incorrect lexical semantics for ``last $n$ days''.
} \fracasex{294	}{\answer yes
\item	Smith was running his own business in two years.
\hyp    Smith ran his own business.
\failsf Parsing error for ``in two years''
} \fracasex{317	}{\answer yes
\item	Every representative has read this report.
\item	No two representatives have read it at the same time.
\item	No representative took less than half a day to read the report.
\item	There are sixteen representatives.
\hyp    It took the representatives more than a week to read the report.
\failsf Counting semantics not implemented
% } \fracasex{318	}{\answer no
% \item	While Jones was updating the program, Mary came in and told him about the board meeting.
% \item	She finished before he did.
% \hyp    Mary's story lasted as long as Jones's updating the program.
% \failsf Incorrect interaction with anaphora
}
  \caption{Selected incorrectly handled examples.
  We list below a number of selected examples from the temporal section
of the FraCaS testsuite where our semantics fails to list the correct
inference. We list the reason why it fails under the item \textbf{F}.
}
  \label{fig:incorrect-examples}
\end{subfigure}
% \caption{Selected examples.}
\end{figure*}
\fi
\section{Results and Evaluation}
\label{sec:results}
Our target is the FraCaS testsuite, which aims at covering a wide
range of common natural-language phenomena. 
%
The suite is structured according to the semantic phenomena involved
in the inference process for each example, and contains nine sections:
Quantifiers, Plurals, Anaphora, Ellipsis, Adjectives, Comparatives,
Temporal, Verbs and Attitudes.  The system described here focuses on
the Temporal section. However, it also supports the other eight
sections.  To our knowledge this is the first system which attempts to
target the temporal section in full. But in fact, our system even provides
support for all the other sections. Thus, a couple of decades after its
formulation, we propose a first attempt at covering the whole suite.
As such, there it is no other system to compare our system with, in
all aspects.  We can however compare with systems which target parts
of the FraCaS testsuite, as shown in \cref{tab:results}.

\providecommand\ncases[1]{{\ensuremath{^{#1}}}}
\begin{table}[hbt]
  \centering
  \small
\begin{tabularx}{\columnwidth}{Xr@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r@{\,\,}r}
Section      & {\kern -2em} \#FraCaS
                          & This        & FC2         & FC & MINE & Nut  & LP  \\ \hline
Quantifiers  & 75         & .93        & .96         & .96    & .77  & .53  & .93  \\
             &            & \ncases{74}& \ncases{74} &        &      &      &     \ncases{44} \\
Plurals      & 33         & .79        & .82         & .76    & .67  & .52  & .73 \\
             &            &            &             &     &   &   & \ncases{24} \\
Anaphora     & 28         & .79        & .86         &   -    & -    & -    &  -       \\
Ellipsis     & 52         & .81        & .87         &   -    & -    & -    &  -       \\
Adjectives   & 22         & .95        & .95         & .95    & .68  & .32  & .73 \\
             &            & \ncases{20}&  \ncases{20}&     &   &   &  \ncases{12} \\
Comparatives & 31         & .65        & .87         & .56    & .48  & .45  &  -       \\
Temporal     & 75         & .73        &  -          &   -    &   -  &  -   &  -       \\
Verbs        & 8          & .75        & .75         &   -    & -    & -    &  -       \\
Attitudes    & 13         & .85        & .92         & .85    & .77  & .46  & .92  \\ 
             &            &            &             &        &      &      & \ncases {9}  \\ \hline
Total        & 337        & .81        & .89         & .83    & .69  & .50  & .85  \\
             &            & \ncases{329}& \ncases{259}& \ncases{174}  & \ncases{174}& \ncases{174}& \ncases{89}
  \end{tabularx}
  \caption{Accuracy of our system compared to others.
    ``This" refers to the approach presented in this paper. When a
    system does not handle the nominal number of test cases (shown in
    the second column), the actual number of test cases attempted is
    shown below the accuracy figure, in smaller font.  ``FC''
    refers to the work of \citet{bernardy_type-theoretical_2017}, and ``FC2'' its followup~\citep{bernardy_wide-coverage_2019}. ``MINE" refers
    to the approach of \citet{Mineshima:2015}, ``NUT" to the CCG
    system that utilises the first-order automated theorem prover
    \textit{nutcracker} \cite{bos:2008}, and ``Langpro"
    to the system presented by \citet{abzianidze_tableau_2015}. A dash
    indicates that no attempt was made for the section. }
  \label{tab:results}
\end{table}
%Additionally, to provide the reader with a better sense of what the
%complete system is capable and not capable of, we list a number of
%correctly and incorrectly handled examples in \cref{fig:correct-examples} and \cref{fig:incorrect-examples}.
\paragraph{Interaction with anaphora}
\label{sec:interaction-with-anaphora}
One reason explaining the lower performance of our system on some
sections of the testsuite is that our interpretation of time interacts
imperfectly with anaphora and ellipsis. Consider the following example:

@@

We have presented a first attempt for a computational approach dealing
with the temporal section of the FraCaS test suite. To do this, we
have provided a simplified taxonomy of aspectual classes for verb
phrases, guided by the applicability of the unicity of action and
temporal subsumption properties. While part of this simplification is
accidental (conflation of activity and accomplishment), we find that
other parts (the automatic distinction between repeatable and
unrepeatable achievements) constitute theoretical improvements.

@@

\cite{bernardy_wide-coverage_2019} presented a
logical system for handling 8 of the 9 sections of the FraCaS test
suite, but excluded section 7, suggesting that it requires many
examples that need an \textit{ad hoc} treatment. Here, we took up this
challenge and have shown that a system similar to theirs can be extended to
cover the remaining section of the test suite, without considerably decreasing the performance of the 
rest of the sections. This is indeed a common problem with logical approaches, namely the fact that one can have theoretically motivated implementations of individual phenomena, e.g. anaphora, ellipsis, quantifiers, temporal reference etc., but when one tries to put all these together into a unified system, this proves to be a daunting task. We believe that this paper presents an exception, and provides a system that can deal with all these different semantic phenomena under a unified system with very good results. 
%
We use the same combination of a number of well-studied tools as \cite{bernardy_wide-coverage_2019} : type
theory, parsing using the Grammatical Framework (GF), Monadic Dynamic
Semantics and proof assistant technology (Coq). The system achieves an
accuracy of 0.73 on the Temporal Section and 0.81 overall.
One of the
things to be looked at is fixing the issues associated with parts of
the test suite that ``broke'' when the temporal analysis was
introduced.  Some of these have been already mentioned: interaction of
the temporal variables with anaphora.

@@

\iffalse\section*{Acknowledgements}
The research reported in this paper was supported by grant 2014-39 from the
Swedish Research Council, which funds the Centre for Linguistic Theory and
Studies in Probability (CLASP) in the Department of Philosophy, Linguistics,
and Theory of Science at the University of Gothenburg. We are grateful to
our colleagues in CLASP for helpful discussion of some of the ideas presented
here. We also thank anonymous reviewers for their useful comments on an
earlier draft of the paper.\fi

\bibliographystyle{acl_natbib}
